### 2014-05-05 rebuttal.tex
Rebuttal letter: general comments and reviewer 3

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

\section{General Comments}

\emph{1. The novelty of this paper with respect to other papers by the same
authors, including one in IJSR 2011, has been questioned. The authors should
convincingly explain what are the novel contributions and the added value of
this work compared to the previous publications.}

We address this question with our reponse to reviewer 3, below.

\emph{2. The form and depth of the presentation are below the standards of an
AIJ article.  The reviewers made many suggestions to make the presentation
better structured, more detailed, and more focused.}

We would like first to thank the reviewers for their thoughtful review of our
work. The manuscript underwent a major rewrite, impacting both the overall style
(which should now be much more pleasant to read) and the whole organisation of
the paper (following the reviewers' recommendations). We also dedicated time to
redraw the figures in an improved academic style.

We address hereafter each of the specific concerns raised by the reviewers (note
that minor typos/style issues have been fixed directly in the manuscript, and
are not discussed here).

\section{Comments from Reviewer 1}

\begin{itemize}

    \item \emph{For longer journal articles, a citation style like 'apalike' that uses the
author names in the references would be more pleasant to read since it saves
reviewers from jumping back and forth all the time.}

As indicated by the Editor in Chief, we need to keep the current citation style.

    \item \emph{p10 l 49: How is this discrimination done?} The discrimination
        algorithms are detailed in \cite{Ros2010b} (Best Paper Award at
        ROMAN2010). We have added a reference to this paper.

    \item \emph{The example scenarios have very bold names for often rather
        simple things.} The reviewer is indeed right. We have mistakenly used internal
        \emph{code names} for the experiments' names. These have been reworded.
\end{itemize}

\section{Comments from Reviewer 2}

\begin{itemize}
    \item \emph{pg 10 (section 3.2.1)
            "The system supports ...". This is one example of a topic that would
            be really interesting to the readers and it is just hinted. In
            particular, learning rules like "cats are animals" maybe really
            nice, but are such rules compatible with the existing knowledge?
        Adding them as triples may work only in absence of contradictions} The
        various natural language grounding processes that are made possible
        within this semantic
        architecture would require a relatively long discussion to be adequately
        explained. They are besides already presented in~\cite{Lemaignan2011a}.
        We have added an explicit reference to this previous publication.

        To specifically answer the reviewer question on how possible
        contradictions are managed, we indeed check that new statements proposed
        by the human do not lead to contradictions before inserting them.

    \item \emph{pg 10 (section 3.2.1)
            "Note that, while the system benefits from the complementary
            modalities, they are not all
            required. The dialogue system can run with only the verbal modality,
            at the cost of a simpler interaction."
        Possibly true, but unjustified claim.} An example has been added to
        clarify this fall-back mechanism.

    \item \emph{"The second level of integration of multi-modality is implicit:
            by computing symbolic properties 
            from the geometry, richer descriptions and hence discrimination
            possibilities are available: 
            for instance, if reachability is available, the robot may ask "do
            you mean the bottle that is 
            accessible to me?" to discriminate between the three bottles. That
            way, procedures relying on 
            discrimination transparently benefit from added modalities." Concept
        remains unclear.} The paragraph has been rephrased and the example
        updated to clarify the concepts.
\end{itemize}

\section{Comments from Reviewer 3}

The main concern of Reviewer 3 relates to the lack of originality of this
contribution.

The reviewer references two previous publications by the authors. \emph{Grounding
the Interaction: Anchoring Situated Discourse in Everyday Human-Robot
Interaction} has indeed been published in 2011 in the \emph{International Journal of
Social Robotics}. However, the overlap with this article appears to be minimal. The
discussion on natural language grounding, which is the focus of the previously
published article, is only briefly summarized in the present article (section
3.3) and as far as we can tell, only one picture, used as an example, as been
reused.

The second publication, \emph{When the robot considers the human...} is supposed
to appear in the proceedings of the \emph{2011 International Symposium on
Robotics Research}, but has yet to be actually pusblished.  While this article
indeed presents our architecture in a way similar to our proposed contribution,
the fact that it belongs to the proceedings of a conference and that it has
actually not yet been published should make it mostly irrelevant.

The reviewer 3 also expresses concerns regarding our proposal not \emph{``helping
researchers to build upon each other's work''} because of the way we
\emph{``presents our design choices''}. We understand this remark has being
related to the comment by Reviewer 2 regarding us not presenting clearly
enough the key issues of human-robot interaction and consequently, the
challenges that HRI brings to artificial intelligence. As explained above, we
have hopefully addressed this concern in the current version of the paper.

Generally speaking, we would like to underline that, while we acknowledge that a
lot of material has been already published on the different sub-systems of
our architecture (as rightfully noted by Reviewer 1), this article give for the
first time a complete picture of the system. This offers, in our opinion, a
perspective on the state of the art in cognitive robotic architectures
for human-robot interaction that may be of interest for the AI and robotics
research communities.

\end{document}